---
title: "Assignment 3"
author: "Inshal Naqvi"
ID: '11721690'
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
Date: 09/20/2021
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Question 1

Before you begin, print thefirst few values of the columns with a headercontaining the string“time”.(head(),head())

```{r read flights.csv}
library(dplyr)
nycflight = read.csv("flights.csv", sep = ",", header = TRUE)

head(select(nycflight, contains("time")))

#summary 
```

#Part a

Count  the number  of flights that  departed  NYC  in  the  first  week  (first  7  days)  of January and February combined. (filter())


```{r filtering -- filter()}
library(dplyr)
filtering = filter(nycflight, month < 3 & day < 8)
nrow(filtering)
```
##Part b

Print the year, month, day, carrier and  air_time of  the flights with  the 6 longest air times, in descending order of air_time. (select(), arrange())

```{r  %>% select() arrang()}
nycflight %>%
  select(year, month, day, carrier, air_time) %>%
  arrange(desc(air_time)) %>%
  head
```
##Part c

Add a new column to the data frame; speed(in miles per hour) is the ratio of distance to air_time. Note that the unit of speed should be miles per hour. If you think they might be useful,  feel  free  to  extract  more  features  than  these,  and  describe  what  they  are. (mutate())


```{r dplyr mutate(adding a col) and speed in miles per hour}
wthspeed=
nycflight %>%
  mutate(speed = distance / (air_time/60)) #airtime divided by 60 to convert
                                            # to miles per hour
head(wthspeed)
```

##Part d

Display the  average,  min  and  max air_time times  for  each month. (group_by(), summarise()). You can exclude NAs for this calculation.

```{r group_by() and Summarise()}
na.omit(nycflight) %>%
  group_by(month) %>%
  summarise(avg_airtime = mean(air_time),
            min_airtime = min(air_time),
            max_airtime = max(air_time))
```

##Part e.1

Impute the missing air_times as the distance divided by the average speed of flights for that destination  (dest).  Make  a  second  copy  of  your  dataframe,  but  this  time  impute missing air_time with  the  average air_time for that destination. What  assumptions  do  these data filling methods make? Which is the bestway to impute the data, or do you see a better way,  and  why?  You  may  impute or  remove other  variables  as  you  find  appropriate. Briefly explain your decisions.(group_by(),mutate())


```{r impute}
wthspeed1=
  nycflight %>%
  mutate(speed = distance / (air_time/60))

newflights1=
wthspeed1 %>%
  
  group_by(dest) %>%
  
  mutate(air_time=ifelse(is.na(air_time), distance / mean(na.omit(speed)), air_time)) %>%
           
           select(air_time, dest, speed)
head(newflights1)

```


##Part e.2


```{r immute_copy}
wthspeed2=

nycflight %>%
  group_by(dest) %>%
    mutate(speed = distance / (air_time/60)) %>%
      mutate(air_time=ifelse(is.na(air_time), mean(na.omit(air_time)),air_time)) %>%
        select(air_time, dest, speed)
head(wthspeed2)

```
The best way in my opinion is by replacing values with mean since that wont have any affect when statistical operation is applied on the value.


##Question2
```{r tidyr}
library(tidyr)
who1<- tidyr::who
who1
```
##Part a
Explain why this line 
 mutate(key=stringr::str_replace(key,"newrel","new_rel"))
is necessary to properly tidy the data. What happens if you skip this line?

#answer
The column "newrel" makes the dataframe inconsistant hence to maintain consistance with the format "newrel" is replaced with "new_rel". This is used later on to extract information in a much cleaner way.


##Part b
How many entries are removed from the dataset when you set values_drop_na to true in the pivot_longer command (in this dataset)?

#answer
First lets check how many entries were there initially.

```{r pivot_longer}
who1 = who %>%
  pivot_longer(
    col = new_sp_m014:newrel_f65,
    names_to = "key",
    values_to = "cases"
  )
head(who1)
```
Here we see that there are 405,440 rows in total. Now lets run pivot_longer and drop the NA values.
```{r pivot no NA}
who1 = who %>%
  pivot_longer(
    col = new_sp_m014:newrel_f65,
    names_to = "key",
    values_to = "cases",
    values_drop_na = TRUE
  )
head(who1)
```
Now we see we that the number of rows are 76,046. This means that rows with NA 329,394


##Part c
Explain the difference between an explicit and implicit missing value, in general. Can you find any implicit missing values in this dataset, if so where?

#answer
Explicit missing values are where the meaning is clearly defined in the place where there is suppose to be a value. Like when there is "NA" wehre there should be a value. With implicit missing values, it is not clearly defined. implicit missing values do not give any clear information. With this data set we dont see any implicitly missing values. One could argue that 0s might be implicitly missing values but in this data set it just means there were no TB cases.

```{r whatever}




```
##Part d
Looking at the features (country, year, var, sex, age, cases) in the tidied data, are they all appropriately typed? Are there any features you think would be better suited as a different type? Why or why not? 

#answer
The year column seems to be good and convey appropriate imformation that it should. However the rest could be represented better by converting them into factor type. Like age should be an int instead of chr.

##Part e
Generatean  informative  visualization,  which shows something  about the  data. Give a brief description of what it shows, and why you thought it would be interesting to investigate.

#Answer
With this we can probably see the speard of Tb with a certain country and check if one variant is spreading more in a certain country

```{r country tb}
library(tidyr) 
who=tidyr::who 
who1 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
who2 <- who1 %>%
mutate(key = stringr::str_replace(key, "newrel", "new_rel"))

who3 <- who2 %>%
separate(key, c("new", "type", "sexage"), sep = "_")
who4 <- who3 %>% select(-new, -iso2, -iso3)
who5 <- who4 %>%
separate(sexage, c("sex", "age"), sep = 1)
who_sp= who5%>% spread(key=country,value=cases)

who_bhutan=
who_sp%>% select(Bhutan,type,year)
Bhutan_TBcases=na.omit(who_bhutan)
head(Bhutan_TBcases)



```
##Part f
Lets construct the table
```{r construct of table}
Group = c(1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3, 3)
Year = c(2006, 2007, 2008, 2009, 2006, 2007, 2008, 2009, 2006, 2007, 2008, 2009)
Qrt.1 = c(15, 12, 22, 10, 12 ,16, 13, 23, 11, 13, 17, 14)
Qrt.2 = c(16, 13, 22, 14, 13, 14, 11, 20,12, 11, 12, 9)
Qrt.3 = c(19, 27, 24, 20, 25, 21, 29, 26, 22, 27, 23, 31)
Qrt.4 = c(17, 23, 20, 16, 18 ,19 , 15, 20, 16, 21, 19, 24)

qrtRev = data.frame(Group, Year, Qrt.1, Qrt.2, Qrt.3, Qrt.4)

head(qrtRev)


```
Lets tidy it up now

```{r gather}
new_qrtRev <- qrtRev %>% gather(Quarter, Revenue, Qrt.1:Qrt.4)
head(new_qrtRev)



```


```{r seperate}
final_qrtRev = new_qrtRev %>% separate(Quarter, c("Time_Interval", "Interval_ID"))
head(final_qrtRev)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
